---
title: "HW8_echo"
format: pdf
author: "Echo (Wenyi) Xu"
editor: visual
embed-resources: true
eval: true
---

Research Question: How do activity categories influence caregivers' moment-to-moment emotions?

Variables - `Activity.Category`: activity that caregivers' were doing when completing the questionnaires: Productive, Social_Face, Social_Virtual, Screen_Time_Leisure, Exercise, Rest_Sleep, Baby_Care - `Composite`: Emotion Composite: Positve Affect, Negative Affect; value from 1 - 5 - `value_wc`: Centered Emotion within and between person

```{r}
# Load the packages
library(dplyr)
library(tidyr)
library(nlme)
library(ggplot2)
library(here)
library(brms)
library(readxl)  # for reading excel files
library(modelsummary)  # for summarizing data
library(cmdstanr)  # use two cores
library(posterior)
library(bayesplot)
library(knitr)

## Data Import
load("/Volumes/data/Studies/STAR/Data/EMA/Saved Environments/data_11_20_2024.RData") 
dfm$Doing <- tolower(dfm$Doing)
```

```{r}
# Data Cleaning 
dfm$Doing <- gsub("e.g.,", "e.g.", dfm$Doing)

# Define groups
productive <- c("work or other productive activities", "chores/errands", 
                "personal maintenance", 
                "transit")
social_face <- c("socializing face-to-face")
social_virtual <- c("socializing virtually")
screen_time_leisure <- c("movies", "using social media") 
exercise <- c("exercising/playing a sport")
rest_sleep <- c("sleeping/napping", "doing absolutely nothing")
baby_care <- c("feeding the baby", "changing the baby")

# Classify activities into categories, accounting for multiple selections, modified including other option, dummy coding each activity 
dfm <- dfm %>%
  mutate(
    Productive = ifelse(grepl(paste(productive, collapse = "|"), Doing, ignore.case = TRUE), 1, 0),
    Social_Face = ifelse(grepl(paste(social_face, collapse = "|"), Doing, ignore.case = TRUE), 1, 0),
    Social_Virtual = ifelse(grepl(paste(social_virtual, collapse = "|"), Doing, ignore.case = TRUE), 1, 0),
    Screen_Time_Leisure = ifelse(grepl(paste(screen_time_leisure, collapse = "|"), Doing, ignore.case = TRUE), 1, 0),
    Exercise = ifelse(grepl(paste(exercise, collapse = "|"), Doing, ignore.case = TRUE), 1, 0),
    Rest_Sleep = ifelse(grepl(paste(rest_sleep, collapse = "|"), Doing, ignore.case = TRUE), 1, 0),
    Baby_Care = ifelse(grepl(paste(baby_care, collapse = "|"), Doing, ignore.case = TRUE), 1, 0)
  ) %>%
  rowwise() %>%
  mutate(
    # Combine all categories into a single string, separated by commas
    Activity.Category = paste(
      na.omit(unique(c(
        ifelse(Productive == 1, "Productive Activities", NA),
        ifelse(Social_Face == 1, "Socializing Face-to-Face", NA),
        ifelse(Social_Virtual == 1, "Socializing Virtually", NA),
        ifelse(Screen_Time_Leisure == 1, "Screen Time (Leisure)", NA),
        ifelse(Exercise == 1, "Exercise", NA),
        ifelse(Rest_Sleep == 1, "Rest and Sleep", NA),
        ifelse(Baby_Care == 1, "Baby Care", NA)
      ))),
      collapse = ", "
    ),
    
    # If more than one category is identified, set as "Other"
    # Activity.Category = ifelse(grepl(",", Activity.Category), "Other", Activity.Category),
    
    # Label as "Other" if no category is identified but `Doing` is not NA
    Activity.Category = ifelse(Activity.Category == "" & !is.na(Doing), "Other", Activity.Category)
  ) %>%
  ungroup()

# Creating a model that aggregates across composites for one per entry 
dfm.ag <- dfm %>% 
  dplyr::group_by(record_id, Composite, Activity.Category, DAY, SIG, Duration..in.seconds.) %>% 
  dplyr::summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE)))%>%  # This is modeling each signal separately
  ungroup()
#####
# Centering within and between person for each emotion
#####
dfm.ag <- dfm.ag %>% group_by(record_id, Composite) %>% 
  dplyr::mutate(w_mean = mean(value, na.rm = T)) %>% ungroup() %>%
  dplyr::mutate(value_bc = scale(value, center = T, scale = F),
                value_wc = value - w_mean) 

dfm.ag <- dfm.ag %>%
  mutate(
    Composite = as.factor(Composite),
    Activity.Category = as.factor(Activity.Category),
    value = as.numeric(value)
  )
dfm.ag <- dfm.ag %>% drop_na(value_wc)
dfm.pos <- dfm.ag %>% filter(Composite == "PosAff")
dfm.neg <- dfm.ag %>% filter(Composite == "NegAff")
dfm.oth <- dfm.ag %>% filter(Composite == "Other")
```

```{r}
## Variable Summary
## plotting
ggplot(
  dfm.ag[!is.na(dfm.ag$Activity.Category) & dfm.ag$Activity.Category != "" & 
         dfm.ag$Composite != "Other" & dfm.ag$Activity.Category != "Mix",], 
  aes(x = Composite, y = value, fill = Activity.Category)
) + 
  my_opts +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = 0.95), alpha = 0.7) +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.3, position = position_dodge(width = 0.95)) +
  labs(
    x = "Emotion", 
    y = "How do you feel right now?"
  ) +
  my_opts + 
  coord_cartesian(ylim = c(1, 3.5)) +
  theme(legend.position = "none")
```

# Model

Let $y$ = value_wc, $g$ = Activity.Category
i represents participant
j represents number of activity catergory
t means observation time points 

Model:
$$
  \begin{aligned}
    y_{ti} & \sim N(\mu_{ti}, \sigma_1) \\
    \mu_{ti}& = \sum_j \beta_{1ij}*g_{ij}+\ \beta_{0i} \\
    \beta_{0i} & \sim N(\mu_{\beta0},\sigma_{\beta0}) \\ 
    \beta_{1ij}& \sim N(\mu_{\beta_{1ij}}, \sigma_{\beta_{1ij}}) \\ 
  \end{aligned}
$$

$y_i$ and $g_{ij}$ are observables.

Prior: 
$$ 
   \begin{aligned} 
      \sigma_{1}& \sim N(0,1) \\
   \sigma_{\beta{0}} & \sim N(0,1)\\
   \mu_{\beta{0}} & \sim N(0,1)\\
   \mu_{\beta_{1ij}}& \sim N(0,1) \\
   \sigma_{\beta_{1ij}}& \sim N(0,1) \\
   \end{aligned} 
$$

```{r}
## Model
#####
# Centering within and between person for each emotion
#####
dfm.ag <- dfm.ag %>% group_by(record_id, Composite) %>% 
  dplyr::mutate(w_mean = mean(value, na.rm = T)) %>% ungroup() %>%
  dplyr::mutate(value_bc = scale(value, center = T, scale = T)[,1],
                value_wc = value - w_mean) 

dfm.ag <- dfm.ag %>%
  mutate(
    Composite = as.factor(Composite),
    Activity.Category = as.factor(Activity.Category),
    value = as.numeric(value)
  )

dfm.ag <- dfm.ag %>% drop_na(value_bc)
dfm.pos <- dfm.ag %>% filter(Composite == "PosAff")
dfm.neg <- dfm.ag %>% filter(Composite == "NegAff")
dfm.oth <- dfm.ag %>% filter(Composite == "Other")

df = dfm.pos

train_size <- floor(0.9 * nrow(df))
train_ind <- sample(seq_len(nrow(df)), size = train_size)
train_df <- df[train_ind, ]
test_df <- df[-train_ind, ]

# Create the matrix `g` from binary indicator columns
# Replace the column names below with the actual names of your binary columns
binary_columns <- c("Productive", "Social_Face", "Social_Virtual", 
                    "Screen_Time_Leisure", "Exercise", "Rest_Sleep", "Baby_Care")

# observations
g <- as.matrix(train_df[binary_columns])
y <- train_df$value_bc
test_g <- as.matrix(test_df[binary_columns])
test_y <- train_df$value_bc
# Number of activities (columns in g)
J <- ncol(g)
# Number of observations (rows in g)
N <- nrow(g)

pid <- factor(train_df$record_id)
P <- length(unique(pid))

# Prepare the data list for Stan
stan_data <- list(
  J = J,      # Number of clusters (activities)
  N = N,      # Number of observations
  y = y,      # Response variable
  g = g,      # Transpose g to match Stan's `J x N` format
  P = P,      # number of participants
  pid = pid   # index of participant
)
```

```{r}
## Results
# Load and compile the Stan model
model <- cmdstan_model("model.stan")
n_chains <- 4

# Fit the model using sampling
fit <- model$sample(
  data = stan_data,        
  chains = n_chains,                 # Number of chains
  parallel_chains = n_chains,        # Run chains in parallel
  iter_sampling = 4000,       # Number of iterations
  iter_warmup = 1000,         # Number of warmup iterations
  thin = 5                    # thinning
)

# Print a summary of the model results
summary <- fit$summary(variables = c("a", "b", "sigma"))

# fit$cmdstan_diagnose()       # Diagnose any issues
fit$draws(variables = c("sigma")) %>% bayesplot::mcmc_trace()  # Trace plot

posterior_predictive <- fit$draws(variables = "ytilde")

# Convert posterior predictive samples to a matrix
y_tilde_matrix <- posterior::as_draws_matrix(posterior_predictive)

# Perform posterior predictive density overlay
bayesplot::ppc_dens_overlay(y[1:125], y_tilde_matrix[,1:125])
```

The results suggest that the model fits okay to the actual data. Futher exploration can focus on simpler or model taking account variables like time points.  
